A Deterministic and Terminating Semantics for the Synchronous Programming Language~\CEU

### 2018-02-08 03-semantics.tex
Small adjusts

The $\NST$ rules are recursive definitions but do not affect the stack level
and never have an emitted event as a precondition:
%
\begin{align*}
\LL p, n,\epsilon \RR &\NST
\LL p',n,e        \RR
    & \textbf{(rule-inner)}
\end{align*}
%
The distinction between $\OUT$ and $\NST$ prevents rules \textbf{push} and
\textbf{pop} to match and inadvertently modify the current stack level before
completing a nested reaction.
%
A complete reaction behaves as follows:
%
\begin{align*}
a) &\5\5
    \LL p,0,ext \RR
        \1\xrightarrow[out]{push}\1
    \LL q,1,\epsilon \RR
\\
b) &\5\5 \1[ \1\xrightarrow[in]{*}\1
    \LL r,i,e \RR
        \1\xrightarrow[out]\1
    \LL s,j,\epsilon \RR \1]*
\\
c) &\5\5 \1\xrightarrow[in]{*}\1
    \LL t,k,\epsilon \RR
        \1\xrightarrow[out]{pop}\1
    \LL u,0,\epsilon \RR
\end{align*}
%
First, a $\xrightarrow[out]{push}$ starts a nested reaction at level $1$ (case
$a$).
Then, a series of alternations between $\NST$ transitions followed by a single
$\OUT$ transition represent nested reactions and stack operations (case $b$).
Finally, at some point, a $\xrightarrow[out]{pop}$ lowers the stack level to
$0$ and terminates the reaction.

The $\NST$ transition rules for the primary expressions are as follows:
%
{ \setlength{\jot}{20pt}
\begin{align*}
\LL mem(id), n, \epsilon \RR &\NST
\LL @nop, n, \epsilon \RR
    & \textbf{(mem)}        \\
%%%
\LL emit(id), n, \epsilon \RR &\NST
\LL @canrun(n), n, id \RR
    & \textbf{(emitInt)}    \\
%%%
\LL @canrun(n), n, \epsilon \RR &\NST
\LL @nop, n, \epsilon \RR
    & \textbf{(canrun)}     \\
\end{align*}
}
%
A $mem$ operation becomes a $nop$ which indicates termination (rule
\textbf{mem}).
An $emitInt(id)$ generates an event $id$ and transits to a $@canrun(n)$ which
can only resume at level $n$ (rule \textbf{emitInt}).
Since nested rules cannot transit with $e \neq \epsilon$, this rule will cause
rule \textbf{push} to execute at the outer level, creating a new level $n+1$ on
the stack.
Also, with the new stack level, the resulting $@canrun(n)$ itself cannot
transit, providing the desired stack-based semantics for internal events.

>>>>

The $\NST$ rules are recursive definitions but do not affect the stack level
and never have an emitted event as a precondition:
%
\begin{align*}
\LL p, n,\epsilon \RR &\NST
\LL p',n,e        \RR
    & \textbf{(rule-inner)}
\end{align*}
%
The distinction between $\OUT$ and $\NST$ prevents rules \textbf{push} and
\textbf{pop} to match and inadvertently modify the current stack level before
completing a nested reaction.
%
A complete reaction behaves as follows:
%
\begin{align*}
a) &\5\5
    \LL p,0,ext \RR
        \1\xrightarrow[out]{push}\1
    \LL q,1,\epsilon \RR
\\
b) &\5\5 \1[ \1\xrightarrow[in]{*}\1
    \LL r,i,e \RR
        \1\xrightarrow[out]\1
    \LL s,j,\epsilon \RR \1]*
\\
c) &\5\5 \1\xrightarrow[in]{*}\1
    \LL t,k,\epsilon \RR
        \1\xrightarrow[out]{pop}\1
    \LL u,0,\epsilon \RR
\end{align*}
%
First (case $a$), a $\xrightarrow[out]{push}$ starts a nested reaction at level
$1$.
Then (case $b$), a series of alternations between $\NST$ transitions and a
single $\OUT$ transition represent, respectively, nested reactions and stack
operations.
Finally (case $c$), at some point, a $\xrightarrow[out]{pop}$ decrements the
stack level to $0$ and terminates the reaction.

The $\NST$ transition rules for primary expressions are as follows:
%
{ \setlength{\jot}{20pt}
\begin{align*}
\LL mem(id), n, \epsilon \RR &\NST
\LL @nop, n, \epsilon \RR
    & \textbf{(mem)}        \\
%%%
\LL emit(id), n, \epsilon \RR &\NST
\LL @canrun(n), n, id \RR
    & \textbf{(emitInt)}    \\
%%%
\LL @canrun(n), n, \epsilon \RR &\NST
\LL @nop, n, \epsilon \RR
    & \textbf{(canrun)}     \\
\end{align*}
}
%
A $mem$ operation becomes a $@nop$ which indicates termination (rule
\textbf{mem}).
An $emitInt(id)$ generates an event $id$ and transits to a $@canrun(n)$ which
can only resume at level $n$ (rule \textbf{emitInt}).
Since nested rules cannot transit with $e \neq \epsilon$, this rule will cause
rule \textbf{push} to execute at the outer level, creating a new level $n+1$ on
the stack.
Also, with the new stack level, the resulting $@canrun(n)$ itself cannot
transit, providing the desired stack-based semantics for internal events.
